## Multiverse analysis in R

Suppose you want to taken each player's position information into account in building the model. As seen in Figure 1, defensive players were more prone to receiving a red card in a match. Controlling for a player's position thus seems like a reasonable choice.

```{r multiverse-tut-position-plot}
tmp <- soccer %>%
  mutate(position = forcats::fct_infreq(position)) %>%
  mutate(position = factor(position, rev(levels(position)))) %>%
  mutate(position_group = forcats::fct_collapse(position,
    Forward = c("Center Forward", "Right Winger", "Left Winger"),
    Midfielder = c("Defensive Midfielder", "Attacking Midfielder", 
                   "Right Midfielder", "Center Midfielder", "Left Midfielder"),
    Defender = c("Center Back", "Right Fullback", "Left Fullback"),
    Goalkeeper = c("Goalkeeper")
  )) %>%
  group_by(playerShort, position, position_group) %>%
  summarise(redCards = sum(redCards), games = sum(games), .groups = "keep")
tmp_group <- tmp %>%
  group_by(position, position_group) %>%
  summarise(medianp = median(redCards / games), n = n(), .groups = "keep")
ggplot(tmp, aes(y = position, x =redCards / games, colour = position_group)) +
  theme_minimal() +
  geom_violin(alpha = 0.25, show.legend = FALSE) +
  geom_point(aes(x = medianp), data = tmp_group, size = 2, shape = 15, show.legend = FALSE) +
  geom_text(aes(x = 0.06, label = paste0(position, " (n=", format(n, big.mark = ",") , ")")),
    size = 3, hjust = 1, vjust = 0.5, nudge_x =- 0.001, show.legend = FALSE,
    data = tmp_group) +
  scale_x_continuous(position = "top", limits = c(0, 0.06)) +
  scale_colour_brewer(palette = "Set2") +
  facet_grid("position_group", scales = "free_y", space = "free_y", switch = "y",
             labeller = as_labeller(~ paste0('<span style="color:',
             RColorBrewer::brewer.pal(4, "Set2")[1:4], ';">',.x, '</span>'))) +
  labs(y = NULL, x = "Proportion of red card received per match", 
       caption = stringr::str_wrap("Figure 1. The violin plots display the distributions of each player's proportion of red cards received in a game and the squares indicate the median value for each position. They suggest that the likelihood of receiving a red card in a soccer match vary by position.", width = 90)) +
  theme(axis.text.y = element_blank(),
        axis.title.x.top = element_text(color = "#555555", size = 8),
        axis.text.x.top = element_text(color = "#555555", size = 8),
        panel.grid.major.y = element_blank(),
        plot.caption = element_text(color = "#555555", size = 10),
        plot.caption.position = "plot",
        strip.text.y.left = ggtext::element_markdown(angle = 0, size = 10))
```

Note that the data set provides _typical position_ each player played _as of June 2014_ when Silberzahn et al. (2018) collected them. However, the red card and game counts may represent an accumulation over multiple years depending on the player's career. During these years, a player may have played multiple positions. This leads to a disconnect between the position information and the red cards on record. For example, a center back as of June 2014 may have received a red card while playing as a right full back. 

To address the issue, one may consider grouping the positions assuming that most players do not switch between the groupings over their careers. Figure 1 shows a common grouping of the positions<sup><a href="#note3">3</a></sup>. We may also consider grouping all positions on the field and only keep _Goalkeeper_ as a separate position. Or, we may even decide not to use the position information deeming the information as unreliable. 

We can argue that all four options below provide _defensible_ analysis paths.

+   Use position variable as is
+   Group positions into the common 4-tie groups as shown in Figure 1
+   Group positions into fielders vs. goalkeeper
+   Do not use position information

### Multiverse analysis using `mverse`

`mverse` is a package that helps define multiverse analyses. We will demonstrate a multiverse analysis across the four defensible analysis paths, or analysis universes. 

First, we start by initiating an `mverse` object with the data set of interest. We will also use the tidyverse package for manipulation.

```{r mverse-tut-init, echo=TRUE, message=FALSE}
library(tidyverse)
library(mverse)
mv <- mverse(soccer) 
```

### Specifying options at decision points

`mverse` specifies different options at each decision point with _branches_. For example, we use `mutate_branch()` to define the options for defining _position_. We can provide each option as an R expression as if we were creating a new column using `mutate()` in tidyverse separated by a comma.

```{r mverse-tut-branch, echo=TRUE}
new_position <- mutate_branch(
  position, # use position as is
  forcats::fct_collapse(
    position, 
    Forward = c("Center Forward", "Right Winger", "Left Winger"), 
    Midfielder = c("Defensive Midfielder", "Attacking Midfielder", 
                   "Right Midfielder", "Center Midfielder", "Left Midfielder"), 
    Defender = c("Center Back", "Right Fullback", "Left Fullback"), 
    Goalkeeper = c("Goalkeeper")), # groups position into 4 groups
  if_else(position == "Goalkeeper", "Goalkeeper", "Fielder") # group all fielders
)
```

To specify the model equation, we use `formula_branch()`. We can provide the model equations as if you were writing the model formula in base R separated by a comma. We need to refer to the name of the branch when including a branched variable. i.e., `new_position`.

```{r mverse-tut-formula, echo=TRUE}
model <- formula_branch(
  cbind(redCards, games - redCards) ~ skintone, # do not use position information
  cbind(redCards, games - redCards) ~ skintone + new_position # include position variable
)
```

To fit the model, we need to add the defined branches to the `mverse` object with data. We will also add a `family_branch()` to specify the underlying distribution of the model. Although we are not creating a branch with multiple options `mverse` requires `family_branch()` to fit logistic regression models. We can add the branches by piping the object through `add_***_branch()` functions<sup><a href="#note4">4</a></sup>.

```{r mverse-tut-add-branches, echo=TRUE}
logit_reg <- family_branch(binomial(link = "logit"))
mv <- mv %>%
  add_mutate_branch(new_position) %>%
  add_formula_branch(model) %>%
  add_family_branch(logit_reg)
```

Note that each branch in `mverse` represents a decision point and by default, all options in each branch are multiplexed throughout all options in other branches. That is, the three options for `position` variable in `mv` are applied for both options of including and excluding the variable in `formula`. To avoid the unnecessary duplicate universes, we can using `branch_condition()` to condition the option without `position` in model equation with one of the options for `position`. When conditioning options, we need to provide the R expressions used to defined the options as shown below.

```{r mverse-tut-branch-condition, echo=TRUE}
condition_position <- branch_condition(
  cbind(redCards, games - redCards) ~ skintone, # formula with no position option
  position # position as is option 
)
```

We then add the branch condition to `mv` option just as we did with branch definitions.

```{r mverse-tut-add-condition, echo=TRUE}
mv <- add_branch_condition(mv, condition_position)
```


### Fitting models across the multiverse

To allow fitting logistic regression models, `mverse` provides `glm_mverse()` which calls `glm()` across the multiverse as defined by the branches. The function executes `glm()` for each unique analysis paths specified by the branches and branch conditions.

```{r mverse-tut-glm, echo=TRUE}
fit <- glm_mverse(mv)
```

### Results

We can use `summary()` function for the multiverse GLM object to extract the estimated coefficients and the corresponding confidence intervals. The output is a tibble that contains estimates for all coefficients across all analysis paths. We can then select the subset that contains estimates for the variable interest, `skintone`. The resulting table consists of 4 rows each representing an analysis path we defined. 

> **Hint:** Turning off the confidence interval with `conf.int = FALSE` can help speed up the process for `glm`.

```{r mverse-tut-glm-summary, echo=TRUE}
est <- summary(fit, conf.int = FALSE) 
filter(est, term == 'skintone')
```

The estimates are in log-odd scale. We can compute the estimates and the confidence intervals in odd scale as below.

```{r mverse-tut-glm-exp, echo=TRUE}
filter(est, term == 'skintone') %>% 
  mutate(
    conf.low = exp(estimate + qnorm(.025) * std.error),
    conf.high = exp(estimate + qnorm(.975) * std.error),
    estimate = exp(estimate)
  ) %>%
  select(universe, estimate, conf.low, conf.high, new_position_branch, model_branch)
```

From the table above, we observe that the results are consistent in direction and statistical significance at 95% confidence level across the multiverse of analysis paths. The results indicate that players with very dark skin tones suffer approximately 30%~35% (Odd ratio 95% CIs ranging approximately within [1.1, 1.6]) higher odds of getting a red card during match compared to their colleagues with very light skin tones. 

```{=html}
<hr />
<footnote id="note3"><sup><a href="#note3">3</a></sup>We referred to <a href="https://yoursoccerhome.com/soccer-positions-a-complete-and-easy-to-understand-guide/" target="_blank">this guide</a> for the groupings.</footnote>
<br /><footnote id="note4"><sup><a href="#note4">4</a></sup><coce>filter_branch()</code> and <code>add_filter_branch()</code> are also available that multiplexes <code>filter()</code> in tidyverse.</footnote>
```
